### 2022-09-20 study manuscript.qmd
ATACH import

---
title: "Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials"
author: "Nikhil Patel, MD"
format: docx
editor: visual
---

## Introduction

Intracerebral hemorrhage encompassess 10-15% of all strokes but is associated with a disproportionate level of death and disability. While recent attention has been paid to avoiding a self-fulfilling prophecy when prognosticating functional outcomes, there is a lingering fear that injuries in certain locations and hemispheres may portend a poorer prognosis.

There is limited research on the outcomes in these situations.

We aimed to conduct a pooled analysis of randomized trials in ICH to examine the following:

-   Is there a difference in functional outcome or HRQOL in patients with left vs. right hemispheric injuries?

## Methods