### 2022-09-20 study protocol.qmd

---
title: "Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials"
subtitle: "Study Protocol and Statistical Analysis Plan"
author: "Nikhil Patel, MD"
format: docx
editor: visual
---

## Introduction

Patients with neurological injury experience different symptoms between the left and right hemispheres, and this may have.

## Methods

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tableone)
library(data.table)
```

### Pooling RCTs

Randomized trials and observational studies included in the analysis:

- ATACH2
- ERICH
- MISTIE
- MISTIE III
- CLEAR-IVH

```{r setup, include=FALSE}

# ATACH2

# Create function to paste together string of FORMxx.txt
atach.import <- function(y) {
  dataset <- paste0("atach2",".",y)
  filepath <- paste0("./ATACH2/FORM",y,".txt")
  dataset <- read.table(filepath, sep = "\t", header = TRUE)
}

atach00 <- atach.import("00")
atach01 <- atach.import("01")
atach02 <- atach.import("02")
atach03 <- atach.import("03")
atach04 <- atach.import("04")
atach05 <- atach.import("05")
# atach05c <- atach.import("05c") Detailed Blood Pressure Data, Unsure if Needed
atach06 <- atach.import("06")
atach07 <- atach.import("07")
atach08 <- atach.import("08")
atach09 <- atach.import("09")
atach10 <- atach.import("10")
atach11 <- atach.import("11")
atach13 <- atach.import("13")
atach14 <- atach.import("14")
atach15 <- atach.import("15")
atach16 <- atach.import("16")
atach17 <- atach.import("17")
atach18 <- atach.import("18")
atach20 <- atach.import("20")
atach21 <- atach.import("21")
atach23 <- atach.import("23")
atach33 <- atach.import("33")

# Merging all tidy data
atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach33)

atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")

# To Do's: write function to quickly import separate files, merge all datasets into one, and extract variables to describe how WLST was done and how often

# https://www.statology.org/merge-multiple-data-frames-in-r/

```