### 2022-09-21 study protocol.qmd
combined atach2 data

---
title: "Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials"
subtitle: "Study Protocol and Statistical Analysis Plan"
author: "Nikhil Patel, MD"
format: docx
---

## Introduction

Patients with neurological injury experience different symptoms between the left and right hemispheres, and this may have.

## Methods

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tableone)
library(data.table)
```

### Pooling RCTs

Randomized trials and observational studies included in the analysis:

-   ATACH2
-   ERICH
-   MISTIE
-   MISTIE III
-   CLEAR-IVH

```{r setup, include=FALSE}

# ATACH2

# Create function to paste together string of FORMxx.txt
atach.import <- function(y) {
  dataset <- paste0("atach2",".",y)
  filepath <- paste0("./ATACH2/FORM",y,".txt")
  dataset <- read.table(filepath, sep = "\t", header = TRUE)
}

atach00 <- atach.import("00")
atach01 <- atach.import("01")
atach02 <- atach.import("02")
atach03 <- atach.import("03")
atach04 <- atach.import("04")
atach05 <- atach.import("05")
# atach05c <- atach.import("05c") Detailed Blood Pressure Data, Unsure if Needed
atach06 <- atach.import("06")
atach07 <- atach.import("07")
atach08 <- atach.import("08")
atach09 <- atach.import("09")
atach10 <- atach.import("10")
atach11 <- atach.import("11")
atach13 <- atach.import("13")
atach14 <- atach.import("14")
atach15 <- atach.import("15")
atach16 <- atach.import("16")
atach17 <- atach.import("17")
atach18 <- atach.import("18")
atach20 <- atach.import("20")
atach21 <- atach.import("21")
atach23 <- atach.import("23")
atach33 <- atach.import("33")
atachct <- read.table("./ATACH2/CTCR.txt", sep = "\t", header = TRUE)

# Merging all tidy data
atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach21, atach33)
atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")
atach.whole <-atach.whole %>% mutate(study = "ATACH2")

# Importing CT Data
atachct <- atachct %>%
  mutate(ctnumber = case_when(
    CTCRMIN < 0 ~ "baseline",
    CTCRMIN > 0 ~ "24hourpost"
  ))

# Extracting First CT Data
atachctbaseline <- atachct %>% 
  filter(ctnumber == "baseline") %>%
  select(SUBJECT_ID, CTCRMIN, CTCRQ05, CTCRQ07, CTCRQ08, CTCRQ09, CTCRQ12, CTCRQ13, CTCRQ11, CTCRQ14) %>%
  rename(CTCRMIN.baseline = CTCRMIN) %>%
  rename(ich.location.baseline = CTCRQ05) %>%
  rename(ivh.baseline = CTCRQ07) %>%
  rename(hydrocephalus.baseline = CTCRQ08) %>%
  rename(pineal.shift.baseline = CTCRQ09) %>%
  rename(ich.volume.baseline = CTCRQ12) %>%
  rename(ivh.volume.baseline = CTCRQ13) %>%
  rename(phe.volume.baseline = CTCRQ11) %>%
  rename(sp.shift.baseline = CTCRQ14)

atachct24hrpost <- atachct %>% # Extracting Second CT Data
  filter(ctnumber == "24hourpost") %>%
    select(SUBJECT_ID, CTCRMIN, CTCRQ05, CTCRQ07, CTCRQ08, CTCRQ09, CTCRQ12, CTCRQ13, CTCRQ11, CTCRQ14) %>%
  rename(CTCRMIN.24hourpost = CTCRMIN) %>%
  rename(ich.location.24hourpost = CTCRQ05) %>%
  rename(ivh.24hourpost = CTCRQ07) %>%
  rename(hydrocephalus.24hourpost = CTCRQ08) %>%
  rename(pineal.shift.24hourpost = CTCRQ09) %>%
  rename(ich.volume.24hourpost = CTCRQ12) %>%
  rename(ivh.volume.24hourpost = CTCRQ13) %>%
  rename(phe.volume.24hourpost = CTCRQ11) %>%
  rename(sp.shift.24hourpost = CTCRQ14)

atach.whole <- left_join(atach.whole, atachctbaseline, by = "SUBJECT_ID")
atach.whole <- left_join(atach.whole, atachct24hrpost, by = "SUBJECT_ID")

# Importing mRS Data

atachmrs30 <- atach09 %>%
  filter(mrsday == 30) %>%
  select(SUBJECT_ID, F09Q01) %>%
  rename(mrs30 = F09Q01)

atachmrs90 <- atach09 %>%
  filter(mrsday == 90) %>%
  select(SUBJECT_ID, F09Q01) %>%
  rename(mrs90 = F09Q01)

atach.whole <-left_join(atach.whole, atachmrs30, by = "SUBJECT_ID")
atach.whole <-left_join(atach.whole, atachmrs90, by = "SUBJECT_ID")

# Table 1

atachVarsTableOne <- c("F33O02", "F01Q01", "F01Q02", "F01Q03M1", "F01Q03M2", "F01Q03M3", "F01Q03M5", "F01Q03M6", "F01Q03M7", "SITE_ID", "COUNTRY", "F03Q01", "F03Q04", "F03Q05", "F03Q06", "F03Q07", "F03Q09", "F03Q12", "F03Q13", "F04Q01", "F04Q03", "F00Q02MIN", "F05Q07a", "F05Q07b", "F33Q04", "F10Q22", "F15Q11", "F15Q04", "F15Q07", "F15Q09", "F15Q12", "F15Q17", "CTCRQ05", "CTCRQ07
", "CTCRQ08", "CTCRQ12", "CTCRQ13", "CTCRQ11", "CTCRQ14", "F07Q01", "F07Q02", "F07Q11", "F07Q04DAY", "F07Q05", "F07Q07", "F07Q08", "F07Q09DAY")

```

Variables included:

-   Age

-   Gender

-   Country Randomized

-   RCT enrolled in

-   Intracerebral hemorrhage

-   Ischemic stroke

-   Coronary artery disease

-   Congestive heart failure

-   Chronic kidney disease

-   Obesity

-   Hypertension

-   Diabetes

-   Antihypertensive therapy

-   Anticoagluation therapy

-   Antiplatelet therapy

-   Lipid lowering therapy

-   Time from onset to presentation

-   Systolic BP

-   Diastolic BP

-   Median GCS

-   Median NIHSS

-   Glucose

-   Hematoma volume

-   Hematoma location

-   IVH

-   ICH Score

-   Extraventricular drain

-   Craniotomy

-   WLST

-   Median time to WLST

-   mRS

-   EuroQOL

-   BI

-   Living at home