### 2023-01-06 manuscript.qmd
manuscript

---
title: "Hemispheric Lateralization and Outcomes After ICH"
subtitle: "A Pooled Analysis of NINDS Studies"
author: "Nikhil Patel, MD"
format: 
  html:
    toc: true
    toc-location: left
    toc-title: ""
    css: styles.css
    anchor-sections: true
    code-fold: true
    code-overflow: wrap
    code-summary: "Show the code"
    output: false
---

# Introduction

Brain injured patients may have different symptoms based on the lesion's laterality

This may affect patient's quality of life or disability down the road

This may also affect clinicians' attitude toward the need for aggressive treatment

We sought to figure out if there was differences in processes among clinicians with left hemispheric versus right hemispheric lesions

We also sought to determine if there are differences in functional outcomes and health-related quality of life

This has never been done in ICH

# Methods

We used data from NINDS clinical trials and 1 observational study: ATACH2, MISTIEII, MISTIEIII, CLEARIII, and ERICH.