### 2023-04-19 code/dataAnalysis/02analysis.qmd
updated code

---
title: "ICH Laterality Analysis"
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
    html-math-method: katex
    css: "styles.css"
---

```{r setup and data import}
#| output: false

library(tidyverse)
library(tableone)
library(data.table)
library(haven) # Importing SAS Files
library(janitor)
library(lubridate)
library(readxl)
library(Hmisc)
library(stringr)
library(qs)
getRs("reptools.r") # Loads reptools.r from Github

d <- qread("./nindsCombined.qs")
```

**Causal question**: does the hemispheric location of an intracerebral hemorrhage (ICH) influence death or major disabilty?

```{r}

d ```