### 2023-04-19 manuscript/manuscript.qmd

---
title: "ICH Laterality Analysis"
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
---

Intracerebral hemorrhage (ICH) is a devastating type of stroke that results from bleeding within the brain tissue. It is a significant cause of morbidity and mortality, with high rates of disability and death. Hemispheric lateralization, which refers to the specialization of the left and right hemispheres of the brain for specific functions, has been suggested to play a role in the functional outcomes of patients with ICH. The left hemisphere is generally associated with language, verbal memory, and logical reasoning, while the right hemisphere is associated with spatial processing, attention, and emotional regulation.

Several studies have investigated the relationship between hemispheric lateralization and functional outcomes in patients with ICH, with conflicting results. Some studies have found that left hemisphere ICH is associated with worse functional outcomes than right hemisphere ICH, while others have found the opposite. Additionally, the specific functions affected by hemispheric lateralization in patients with ICH remain unclear.

Understanding the relationship between hemispheric lateralization and functional outcomes in patients with ICH is essential for improving patient care and developing targeted rehabilitation strategies. Therefore, this paper aims to review the current literature on hemispheric lateralization and functional outcomes in patients with ICH. We will discuss the potential mechanisms underlying these relationships and highlight the implications for clinical practice.