### 2023-08-21 code/dataAnalysis/logistic.qmd
who knows

---
title: "Hemispheric Lateralization in ICH"
subtitle: "Logistic Regression"
author: "Nikhil Patel"
project:
  type: default
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
    html-math-method: katex
    css: "styles.css"
---

```{r hemisphere only}
priors <- c(prior())

```