### 2023-09-29 analysis/clean.qmd
re-arranging

---
title: "Process and Clean Data"
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(targets)
library(scales)
library(patchwork)
library(DT)
library(naniar)
library(here)

```

*Data has been cleaned and merged elsewhere.*

```{r description of data}

data <- read_csv(here("data", "all.csv"))

```