### 2023-09-29 analysis/dag.qmd

---
title: "Causal Assumptions - DAG"
---

```{r setup}

library(tidyverse)
library(dagitty)
library(ggdag)

```

```{r dag}

dag_laterality <- dagify(ich_laterality ~ functional_outcome)

dag_laterality |>
  ggdag()

```