### 2023-09-29 analysis/dag.qmd
dags, bruh

---
title: "Causal Assumptions - DAG"
code-fold: true
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(dagitty)
library(ggdag)

```

## Specify a Causal Question

```{r dag}
#| fig-width: 10

dag_laterality <- dagify(ich_laterality ~ functional_outcome)

dag_laterality |>
  ggdag()

```