### 2023-10-02 analysis/descriptive.qmd
more frustration

#
#
#
#
#

library(tidyverse)
library(dagitty)
library(ggdag)
library(MetBrewer)
library(here)
library(tableone)
library(gt)
library(gtExtras)
library(broom)

data <- read_csv(here("data", "all.csv"))

#
#
#
#
#
#| fig-width: 20

# have to fix time from symptom onset to presentation in ED (units seems to be jacked up)

table1_vars <- c("age", "sex", "race", "ethnicity", "htn", "dm2", "tobacco", "sbp_baseline", "gcs_baseline", "nihss_baseline", "ich_volume_baseline", "ivh", "ich_location", "neurosurgery_evac", "comfort_care_binary", "study")

table1 <- CreateTableOne(vars = table1_vars, data = data, strata = "ich_laterality")

table1

#
#
#