### 2023-10-02 analysis/models.qmd

---
title: "Models"
code-fold: true
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(marginaleffects)
library(brms)
library(ggplot2)
library(ggdist)
library(broom)
library(here)

data <- read_csv(here("data", "all.csv"))

```

## Estimate the Causal Effect

First let's do a replication of the INTERACT re-analysis.

```{r interact2 analysis replication}

priors <- c(prior(normal(0, 1.5), class = Intercept))

inits <- list(Intercept = c(-2, -1, 0, 1, 2, 2.5))

# mod <- brm(mrs_90 ~ 1,
#   data = data,
#   family = cumulative("logit"),
#   prior = priors,
#   init = rep(list(inits), 4),
#   chains = 4, iter = 5000 * 2, seed = 28203, cores = 4,
#   backend = "cmdstanr", refresh = 0
# )

```

## Conduct Sensitivity Analyses