### 2023-10-03 analysis/logistic.qmd
logistic

---
title: "Logistic Models"
code-fold: true
---

These analyses are intended to build up intuition on using Bayesian logistic regression as a stepping stone to understanding and using Bayesian ordinal regression. We will work with dichotimized variables of the modified Rankin score as well as the EuroQOL.

```{r setup}
#| code-fold: true
#| code-summary: "Setup"
#| warning: false
#| message: false

library(bayesrules)
library(tidyverse)
library(janitor)
library(brms)
library(cmdstanr)
library(rstanarm)
library(marginaleffects)
library(broom)
library(broom.mixed)
library(parameters)
library(tidybayes)
library(ggdist)
library(patchwork)
library(latex2exp)
library(scales)
library(here)

# plot stuff
colors <- MetBrewer::met.brewer("Cross", 6)
theme_set(theme_bw(base_family = "Cantarell"))
bayesplot::color_scheme_set(colors)

# seeds
set.seed(160)
BAYES_SEED <- 160

# data
data <- read_rds(here("data", "all.rds"), refhook = NULL)

data <- data |>
  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
  filter(study == "ERICH" | study == "ATACH2") |>
  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04) |>
  drop_na(ich_laterality)

data$ich_location <- fct_drop(data$ich_location)

```

## General Setup

We are interested in modeling the probability of a "good" outcome for patients with ICH at 90 days as defined by the **modified Rankin Scale**. For this session, we will work with a dichotimized version of the mRS at various cutpoints (0-1 vs 2-6, 0-2 vs 3-6, 0-3 vs 4-6, 0-45 vs 5-6). All of these cutpoints have been used in various randomized trials and observational studies.

We will base our model on these predictors:

- `age`
- `ich_laterality`: the main "predictor" of interest for the main analysis
- `ich_location`: only supratentorial locations are used (lobar, thalamus, basal ganglia)
- `ich_volume_baseline`
- `ivh`
- `neurosurgery_evac`

### Modeling poor outcome given age and ICH Volume

We will use continuous variables first before jumping into categorical predictors.

```{r}
#| fig-width: 10

data |>
  drop_na(mrs_90_03) |>
  ggplot(aes(x = mrs_90_03)) +
  geom_density(fill = colors[3], alpha = 0.2) +
  facet_wrap(~ich_location)

data |>
  drop_na(mrs_90_02) |>
  tabyl(ich_location, mrs_90_02) |>
  adorn_percentages()

```