### 2023-10-18 analysis/eda.qmd

---
title: "Exploratory Data Analysis"
code-fold: true
---

{{< include 00-setup.qmd >}}

## Outcomes

```{r aggressiveness, warning=FALSE, message=FALSE}
#| fig-width: 10

```

```{r euroqol, warning=FALSE, message=FALSE}
#| fig-width: 10

data <- read_csv(here("data", "all.csv")) |>
  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
  filter(!is.na(ich_laterality))

data |>
  ggplot(aes(euro_vas_90, fill = ich_laterality)) +
  geom_density(color = NA, alpha = .8)

```