### 2024-01-12 analysis/04-aggressive_care.qmd

---
title: "Differences in Care"
code-fold: false
toc-depth: 5
---

{{< include 00-setup.qmd >}}

```{r color setup, warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: Setup

# Left/Right hemisphere colors
clrs_saguaro <- NatParksPalettes::natparks.pals("Saguaro")
clr_left <- clrs_saguaro[1]
clr_right <- clrs_saguaro[6]
clr_diff <- clrs_saguaro[4]

```

**How are patients treated differently if they experience a left hemispheric ICH vs. right hemispheric ICH?**

- Do they receive neurosurgical intervention at a different rate?
- Are they on the ventilator for longer?
- Are they less likely to receive a tracheostomy or PEG?
- Are they more likely to proceed to comfort care?

There are five studies available to analyze: `ERICH`, `ATACH-2`, `MISTIE-II`, `CLEAR-IVH`, and `MISTIE-III`. Since we are interested in the potential differences in rates of neurosurgical intervention, we will exclude studies that randomized patients to an intervention, leaving `ERICH` and `ATACH-2` to help answer these questions.

### Unadjusted Analysis

::: {.callout-warning appearance="simple"}

Things to clarify:

- How should `comfort care` be defined? Below, it is a binary variable, but would it be better to model it based on a time-to-event process?
- Run "unadjusted analyses" and put results of both adjusted and unadjusted into one table?

:::

```{r table two unadjusted, warning=FALSE, message=FALSE}

table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")

table2 <- data_erich_atach |>
  select(table_two_vars) |>
  tbl_summary(
    by = ich_laterality,
    missing = "no",
    label = list(
      neurosurgery_evac ~ "Neurosurgical Intervention",
      evd ~ "Ventriculostomy",
      days_mechanical_ventilation ~ "Mechanical Ventilation Days",
      tracheostomy ~ "Tracheostomy",
      comfort_care_binary ~ "Comfort Care"
    )
  ) |>
  add_overall() |>
  bold_labels()

table2

```

It appears that patients with left hemispheric lesions tend to receive neurosurgical intervention at a lower rate, as well as stay on the ventilator for longer (based on IQR). There does not seem to be a large difference in rates of tracheostomy and comfort care.

### Modeling Differences in Care

#### Neurosurgical Intervention

> How does the probability of neurosurgical intervention (outcome) change if a patient has a left hemispheric ICH vs. right hemispheric ICH?

We will incoporate the following predictors into the model: `ich_laterality`, `age`, `ich_location`, `gcs_baseline`, `ich_volume_baseline`, `ivh`, and `time_symptoms_to_ed`.

##### Priors

We are using default priors as defined in the `brms` package.

::: {.callout-warning appearance="simple"}

Think about defining weakly informative priors and running prior predictive checks to make this analysis more robust.

:::

##### Posterior simulation

```{r neurosurgical intervention, warning=FALSE, message=FALSE}
#| fig-width: 10

neurosurgery_glm <- brm(
  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = bernoulli(link = "logit")
)

neurosurgery_glm |>
  gather_draws(`^b_.*`, regex = TRUE) |>
  ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
  geom_line(size = 0.1) +
  scale_color_viridis_d(option = "rocket", end = 0.85) +
  facet_wrap(vars(.variable), scales = "free_y")

pp_check(neurosurgery_glm, ndraws = 500)

model_parameters(neurosurgery_glm,
  centrality = "mean", dispersion = TRUE,
  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
  exponentiate = TRUE
) |>
  print(digits = 2, zap_small = TRUE)

```

Based on this model, there is a **much higher odds (~120% increase)** of a patient with right hemispheric ICH receiving neurosurgical intervention as compared to a patient with left hemispheric ICH **(OR 2.24, 95% CI 1.55 - 3.14)**.

```{r neurosurgical intervention marginaleffects, warning=FALSE, message=FALSE}
#| fig-width: 10

```