### 2024-02-13 reports/aggressiveness.qmd
dags

---
title: Differences in care for patients experiencing left vs. right hemisphere ICH
author: Nikhil Patel, MD
format: html
toc: true
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
#| code-fold: false

library(targets)
library(readr)
library(tidyr)
library(skimr)
library(dplyr)
library(gt)
library(gtsummary)
library(ggplot2)
library(ggdag)
library(dagitty)
library(marginaleffects)
library(modelsummary)

theme_set(theme_minimal(base_family = "Figtree"))

# tar_load(ich)
# tar_load(non_surgery_trials)
# tar_load(surgery_trials)
# tar_load(table_1)
# tar_load(table_2)
# tar_load(aggressive_dag)
tar_load_everything()

```

## Exploratory Data Analysis

### Data Overview

::: {.callout-warning title="Issues to Address"}
`htn` is missing mostly all information. It is also listed as a character variable.

All other co-morbidities are missing information; if this is the case, should default to "No".

`ivh` is also missing some values, which should also default to "No"?

Any "Yes/No" variable should not have any missing information, should just default to "No"

Verify accuracy/validity of `time_symptoms_to_ed`, `sbp_baseline`,

Fix/drop unused levels in `race`, and fix "American Indian"

Fix "unknown" category in `ethnicity` to "Not Hispanic or Latino"

Fix/drop unused levels in `ich_location` and `study`
:::

```{r}

glimpse(non_surgery_trials)
skim(non_surgery_trials)

```

### Table 1

::: {.callout-warning title="Issues to Address"}

Collapse unused levels in `race`, `ethnicity`, `ich_location` and `study`

Change order of the displayed factors to frequency

:::

```{r}

table_1

```

### Relationships between Variables

```{r}
#| fig-align: center

table_2

```

## Statistical Models

::: {.callout-important title="Other Considerations"}

Add analyses stratifying by each study: `ERICH` and `ATACH2`

Construct DAGs for **each** outcome (e.g., each model)!

Add `DNR` in addition to `comfort_care`?

:::

* `neurosurgery_evac`: logistic
* `ventriculostomy`: logistic
* `days_mechanical_ventilation`: poisson (cumulative logit?)
* `tracheostomy`: logistic
* `comfort_care`: logistic (days to comfort care, time to event?)

### DAGs

```{r}
#| width: 1200
#| height: 2000
#| fig-align: center
#|
theme_set(
  theme_dag() %+replace%
    # also add some additional styling
    theme(
      legend.position = "bottom",
      strip.text.x = element_text(margin = margin(2, 0, 2, 0, "mm"))
    )
)

ggdag_adjustment_set(aggressive_dag, shadow = TRUE, use_labels = "label", text = FALSE)

ggdag_adjustment_set(outcome_dag, shadow = TRUE, use_labels = "label", text = FALSE)

aggressive_dag |>
  ggdag_paths(shadow = TRUE, text = FALSE, use_labels = "label")

```