### 2024-02-15 reports/outcomes.qmd

---
title: Differences in outcomes for patients experiencing left vs. right hemisphere ICH
author: Nikhil Patel, MD
format:
  html: 
    mainfont: Figtree
    monofont: Source Code Pro
    fontsize: 0.97rem
    embed-resources: true
    toc: true
    toc-location: left
    toc-depth: 6
    toc-expand: true
    toc-title: Contents
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
    highlight-style: nord
---

::: {.callout-note icon="false"}
## Research question

Do patients presenting with left hemispheric ICH have different functional outcomes (as measured by mRs and EuroQoL at 90 days) as compared to those with right hemispheric ICH?

:::

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
library(targets)
library(skimr)
library(dplyr)
library(gt)
library(gtsummary)
library(ggplot2)
library(ggdag)
library(dagitty)
library(marginaleffects)
library(modelsummary)
library(MetBrewer)

# theme customization
theme_set(theme_minimal(base_family = "Figtree"))

theme_set(
  theme_dag(base_family = "Figtree") %+replace%
    # also add some additional styling
    theme(
      legend.position = "bottom",
      strip.text.x = element_text(margin = margin(2, 0, 2, 0, "mm"))
    )
)

status_colors <- c(exposure = "#CE4441", outcome = "#004F63", latent = "grey80")

# load all targets objects
# tar_load_everything()
tar_load(table_1_outcomes)
tar_load(dag_outcomes)
tar_load(ich_all)

```

## Exploratory Data Analysis

### Data Overview

::: {.callout-warning title="Issues to Address"}

?

:::

```{r}

glimpse(ich_all)
skim(ich_all)

```

### Table 1

::: {.callout-warning title="Issues to Address"}

?

:::

```{r}

table_1_outcomes

```

### Relationships between Variables

::: {.callout-important title="Other Considerations"}

?

:::

## Statistical Models

::: {.callout-important title="Other Considerations"}

Add analyses stratifying by each study: `ERICH` and `ATACH2`

?

:::

### Outcomes

We aim to estimate the total direct effect of hemispheric laterality in intracerebral hemorrhage on the probability of a good functional outcome, as defined by the following:

* `mrs_90`: modified Rankin at 90 days: cumulative logit
* `euro_vas_90`: EuroQOL VAS at 90 days: beta regression
* `euro_90`: all 5 domains of the EuroQOL at 90 days: cumulative logit

#### DAG

```{r outcomes dag}
#| fig-align: center
#| fig.height: 9

dag_outcomes |>
  tidy_dagitty() |>
  node_status() |>
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(aes(color = status)) +
  geom_dag_edges_diagonal() +
  geom_dag_label_repel(aes(label = label, fill = status),
    seed = 51,
    color = "white", fontface = "bold", size = 4.5
  ) +
  scale_color_manual(values = status_colors, na.value = "grey20") +
  scale_fill_manual(values = status_colors, na.value = "grey20")

```

#### DAG-Implied Adjustment Sets

Similar to above, there are no

```{r outcomes dag adjustment sets}
#| code-fold: false
#| fig-align: center
#| fig.height: 10
#| fig.width: 20

adjustmentSets(
  dag_outcomes,
  type = "minimal",
  effect = "direct",
  max.results = Inf
)

adjustmentSets(
  dag_outcomes,
  type = "canonical",
  effect = "total",
  max.results = Inf
)

```