### 2024-02-20 reports/posteriors.qmd
cumulative logit woes

---
title: Differences in care for patients experiencing left vs. right hemisphere ICH
subtitle: Posteriors!
author: Nikhil Patel, MD
format:
  html: 
    mainfont: Figtree
    monofont: Source Code Pro
    embed-resources: true
    toc: true
    toc-location: left
    toc-depth: 6
    toc-title: Contents
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
    code-link: true
    highlight-style: nord
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
library(targets)
library(tidyverse)
library(glue)
library(marginaleffects)
library(modelsummary)
library(ggdist)
library(gghalves)
library(scales)
library(systemfonts)
library(brms)
library(tidybayes)

# theme customization
theme_set(theme_minimal(base_family = "Figtree"))

# Format things as percentage points
label_pp <- label_number(
  accuracy = 1, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

label_pp_tiny <- label_number(
  accuracy = 0.01, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

# load targets objects
tar_load(model_parameters)
tar_load(posterior_neurosurgery_minimal)
tar_load(posterior_neurosurgery_canonical)
tar_load(posterior_evd_minimal)
tar_load(posterior_evd_canonical)
tar_load(posterior_trach_minimal)
tar_load(posterior_trach_canonical)
tar_load(posterior_vent_minimal)
tar_load(posterior_vent_canonical)
tar_load(posterior_comfort_care_minimal)
tar_load(posterior_comfort_care_canonical)
tar_load(posterior_early_wlst_minimal)
tar_load(posterior_early_wlst_canonical)
tar_load(posterior_dnr_minimal)
tar_load(posterior_dnr_canonical)

# posterior functions

posterior_printing <- function(x, y, z) {
  p1 <- model_parameters$priors[x]

  p2 <- y

  p3 <- predictions(y, by = "ich_laterality") |>
    as_tibble()

  p35 <- avg_comparisons(y, variables = "ich_laterality")

  p4 <- avg_comparisons(y, variables = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw)) +
    stat_halfeye(
      .width = c(0.8, 0.95), point_interval = "median_hdi",
      fill = "#bc3032"
    ) +
    scale_x_continuous(labels = label_pp) +
    labs(
      title = "Posterior Prediction",
      x = paste("Effect of hemispheric laterality\n on the probability of", {{ z }}),
      y = NULL, caption = "80% and 95% credible intervals shown in black"
    )

  p5 <- y |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    ggplot(aes(ich_right_or)) +
    stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "bc3032")) +
    theme(legend.position = "none")

  p6 <- y |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      or_1 = sum(ich_right_or > 1) / n(),
      or_1.1 = sum(ich_right_or > 1.1) / n(),
      or_1.2 = sum(ich_right_or > 1.2) / n()
    )

  p5 <- pp_check(p2, ndraws = 500)

  return(list(
    priors = p1,
    model = p2,
    fit = p5,
    predictions = p3,
    comparison = p35,
    plot = p4,
    plot2 = p5,
    probabilities = p6
  ))
}

```

## Posterior Predictions

### Neurosurgical Intervention

#### Minimal Model

```{r}
posterior_printing("neurosurgery_minimal", posterior_neurosurgery_minimal, "receiving neurosurgical intervention")

```

#### Canonical Model

```{r}

posterior_printing("neurosurgery_canonical", posterior_neurosurgery_canonical, "receiving neurosurgical intervention")

```

### EVD

#### Minimal Model

```{r}

posterior_printing("evd_minimal", posterior_evd_minimal, "receiving an EVD")

```

#### Canonical Model

```{r}

posterior_printing("evd_canonical", posterior_evd_canonical, "receiving an EVD")

```

### Tracheostomy

#### Minimal Model

```{r}

posterior_printing("trach_minimal", posterior_trach_minimal, "receiving an tracheostomy")

```

#### Canonical Model

```{r}

posterior_printing("trach_canonical", posterior_trach_canonical, "receiving a tracheostomy")

```

### Days of Mechanical Ventilation

#### Minimal Model

```{r}

posterior_printing("vent_minimal", posterior_vent_minimal, "number of days of mechanical ventilation")

```

#### Canonical Model

```{r}

posterior_printing("vent_canonical", posterior_vent_canonical, "number of days of mechanical ventilation")

```

### Comfort Care

#### Minimal Model

```{r}

posterior_printing("comfort_care_minimal", posterior_comfort_care_minimal, "comfort care")

```

#### Canonical Model

```{r}
posterior_printing("comfort_care_canonical", posterior_comfort_care_canonical, "comfort care")

```

### Early WLST

#### Minimal Model

```{r}
posterior_printing("early_wlst_minimal", posterior_early_wlst_minimal, "early WLST")

```

#### Canonical Model

```{r}
posterior_printing("early_wlst_canonical", posterior_early_wlst_canonical, "early WLST")

```

### DNR

#### Minimal Model

```{r}
posterior_printing("dnr_minimal", posterior_dnr_minimal, "DNR")

```

#### Canonical Model

```{r}
posterior_printing("dnr_canonical", posterior_dnr_canonical, "DNR")

```