### 2024-02-20 reports/posteriors.qmd

---
title: Differences in outcomes for patients experiencing left vs. right hemisphere ICH
author: Nikhil Patel, MD
format:
  html
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}

library(targets)
library(tidyverse)
library(glue)
library(marginaleffects)
library(modelsummary)
library(ggdist)
library(gghalves)
library(scales)
library(systemfonts)
library(brms)
library(tidybayes)

# theme customization
theme_set(theme_minimal(base_family = "Figtree"))

# Format things as percentage points
label_pp <- label_number(
  accuracy = 1, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

label_pp_tiny <- label_number(
  accuracy = 0.01, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

# load targets objects
tar_load(model_parameters)
tar_load(posterior_neurosurgery_minimal)
tar_load(posterior_neurosurgery_canonical)
tar_load(posterior_evd_minimal)
tar_load(posterior_evd_canonical)
tar_load(posterior_trach_minimal)
tar_load(posterior_trach_canonical)
tar_load(posterior_vent_minimal)
tar_load(posterior_vent_canonical)
tar_load(posterior_comfort_care_minimal)
tar_load(posterior_comfort_care_canonical)
tar_load(posterior_early_wlst_minimal)
tar_load(posterior_early_wlst_canonical)
tar_load(posterior_dnr_minimal)
tar_load(posterior_dnr_canonical)

```