### 2024-07-08 manuscripts/manuscript1_supplement.qmd

---
title: Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention
author: "Nikhil Patel, MD"
subtitle: Supplementary Appendix
authors:
  - name: Nikhil Patel
    orcid: 0000-0002-3193-8590
    affiliation: Department of Internal Medicine, Wake Forest University School of Medicine
    roles: data analysis, writing
    corresponding: true
  - name: Matt Jaffa
    affiliation: Department of Neurology, University of Connecticut School of Medicine
  - name: Rahul Karamchandani
    affiliation: Department of Neurology, Wake Forest University School of Medicine
  - name: Andrew Asimos
    affiliation: Department of Emergency Medicine, Wake Forest University School of Medicine
  - name: Jonathan Parish
    affiliation: Carolina Neurosurgery & Spine Associates
  - name: Nicholas Allen Morris
    affiliation: Department of Neurology, University of Maryland School of Medicine
date: last-modified
abstract: |

table-of-contents: true
toc-title: Contents
---

```{r setup}
#| echo: false
#| output: false

#| echo: false
#| fig-height: 12

library(targets)
library(tidyverse)
library(withr)
library(here)
library(modelsummary)
library(marginaleffects)
library(brms)
library(tidybayes)
library(bayesplot)
library(glue)
library(ggdist)
library(patchwork)
library(tinytable)
library(gt)

# plot stuff
theme_ich <- function(base_family = "Figtree",
                      base_size = 10) {
  theme_minimal(base_size = base_size) +
    theme(
      text = element_text(lineheight = 2),
      plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
      plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
      axis.text.y = element_text(size = rel(0.8)),
      axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
      axis.text.x = element_text(size = rel(2)),
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.justification = 1,
      panel.grid = element_line(colour = "#F3F4F5"),
      plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
    )
}

# load data and models
with_dir(here(), tar_load(ich_aggressive))

```

# Details of Statistical Analysis

## Covariate Selection

```{r dags}
#| echo: false

# dag_aggressive

```

## Prior Distribution Definitions

```{r prior definitions}
#| echo: false
#| output: false

neutral_mean <- 0
neutral_sd <- 0.35
neutral_lower_OR <- exp(neutral_mean - (1.96 * neutral_sd))
neutral_upper_OR <- exp(neutral_mean + (1.96 * neutral_sd))

left_mean <- -0.22
left_sd <- 0.175
left_lower_OR <- exp(left_mean - (1.96 * left_sd))
left_upper_OR <- exp(left_mean + (1.96 * left_sd))

right_mean <- 0.18
right_sd <- 0.175
right_lower_OR <- exp(right_mean - (1.96 * right_sd))
right_upper_OR <- exp(right_mean + (1.96 * right_sd))

```

In defining prior distributions, we generally followed suggestions from previous literature. [@zampieriUsingBayesianMethods2021; @kruschkeBayesianAnalysisReporting2021; @gelmanBayesianWorkflow2020] Specifically, we defined 4 sets of priors for each outcome:

**Neutral prior**: mean of **`r neutral_mean`** and standard deviation of **`r neutral_sd`**, centering around an OR of **`r exp(neutral_mean)`** and 95% of the the probability mass for the OR between **`r round(neutral_lower_OR, digits = 2)`** to **`r round(neutral_upper_OR, digits = 2)`**.

**Left hemisphere prior**: mean of **`r left_mean`** and standard deviation of **`r left_sd`**, centering around an OR of **`r round(exp(left_mean), digits = 2)`** and 95% of the the probability mass for the OR between **`r round(left_lower_OR, digits = 2)`** to **`r round(left_upper_OR, digits = 2)`**.

**Right hemisphere prior**: mean of **`r right_mean`** and standard deviation of **`r right_sd`**, centering around an OR of **`r round(exp(right_mean), digits =2)`** and 95% of the the probability mass for the OR between **`r round(right_lower_OR, digits = 2)`** to **`r round(right_upper_OR, digits =2)`**.

**Flat prior**: most mimics a frequentist analysis

```{r prior graphical summary}
#| echo: false
#| fig-height: 6

# ggplot(data = tibble(x = -2:2), aes(x = x)) +
#   geom_function(fun = dnorm, args = list(mean = -20, sd = 0.5), linewidth = 1) +
#   # stat_function(fun = dnorm_limit, geom = "area", color = "orange", alpha = 0.2) +
#   geom_vline(xintercept = -0.22, linetype = "dashed", color = "orange", linewidth = 1) +
#   scale_x_continuous(breaks = seq(-2, 2, 0.5)) +
#   theme_ich()

```

### Prior Predictive Checks

```{r prior predictive check - function}
#| echo: false
#| output: false
## | # | | fig-height: 12

f_prior_predictive_check <- function(neutral_model, left_model, right_model, flat_model) {
  models <- list(
    "Neutral Prior" = neutral_model,
    "Left Hemisphere Prior" = left_model,
    "Right Hemisphere Prior" = right_model,
    "Flat Prior" = flat_model
  )

  # gather summary variables for each model
  neutral <- models$"Neutral Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  left <- models$"Left Hemisphere Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or < 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  right <- models$"Right Hemisphere Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  flat <- models$"Flat Prior" |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      aor = quantile(ich_right_or, 0.5),
      lower_ci = quantile(ich_right_or, 0.025),
      upper_ci = quantile(ich_right_or, 0.975),
      or_1 = sum(ich_right_or > 1) / n(),
      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
    )

  prior_summary <- rbind(neutral, left, right, flat) |>
    mutate(prior = c("Neutral", "Left Hemisphere", "Right Hemisphere", "Flat")) |>
    select(c(prior, aor, lower_ci, upper_ci, or_1, rope))

  # add prior predictions
  neutral_predictions <-
    as_tibble(avg_predictions(models$"Neutral Prior", by = "ich_laterality")) |>
    mutate("prior" = "Neutral") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  left_predictions <-
    as_tibble(avg_predictions(models$"Left Hemisphere Prior", by = "ich_laterality")) |>
    mutate("prior" = "Left Hemisphere") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  right_predictions <-
    as_tibble(avg_predictions(models$"Right Hemisphere Prior", by = "ich_laterality")) |>
    mutate("prior" = "Right Hemisphere") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  flat_predictions <-
    as_tibble(avg_predictions(models$"Flat Prior", by = "ich_laterality")) |>
    mutate("prior" = "Flat") |>
    select(c("prior", ich_laterality, estimate)) |>
    pivot_wider(names_from = ich_laterality, values_from = estimate)

  prediction_summary <- rbind(neutral_predictions, left_predictions, right_predictions, flat_predictions)

  prior_prediction_summary <- left_join(prediction_summary, prior_summary, by = "prior") |>
    gt(rowname_col = "prior") |>
    tab_stubhead(label = "Type of Prior") |>
    fmt_number(columns = everything(), decimals = 2) |>
    fmt_number(columns = 6, rows = 4, decimals = 0) |>
    fmt_percent(columns = 2:3, rows = 1:4, decimals = 1) |>
    tab_header(
      title = md("**Prior Predictive Checks**")
    ) |>
    cols_label(
      Left = "Left Hemisphere",
      Right = "Right Hemisphere",
      aor = "aOR",
      lower_ci = "Lower 95% Credible Interval",
      upper_ci = "Upper 95% Credible Interval",
      or_1 = "Probability of Difference (aOR >1)",
      rope = "% of Distribution outside of ROPE"
    ) |>
    cols_width(
      1 ~ pct(50)
    ) |>
    tab_footnote(
      footnote = "For left hemisphere prior, the p(aOR) < 1 is shown",
      locations = cells_column_labels(columns = 7)
    )

  return(prior_prediction_summary)
}

results_visual <- function(model) {
  model <- model

  draws <- model |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize()

  pred <- predictions(model, by = "ich_laterality") |>
    tibble()

  theme_ich <- function(base_family = "Inter",
                        base_size = 14) {
    theme_minimal(base_size = base_size) +
      theme(
        text = element_text(lineheight = 2),
        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
        axis.text.y = element_text(size = rel(0.8)),
        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
        axis.text.x = element_text(size = rel(2)),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.justification = 1,
        panel.grid = element_line(colour = "#F3F4F5"),
        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
      )
  }

  p1 <- predictions(model, by = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw, fill = ich_laterality)) +
    geom_density(alpha = .7) +
    scale_fill_manual(values = c("#054224", "#ff8800")) +
    scale_x_continuous(labels = scales::percent) +
    labs(
      title = "Predicted Probability",
      fill = "ICH Laterality",
      y = NULL
    ) +
    theme_ich()

  return(p1)
}

```

#### Neurosurgical Intervention

```{r prior predictive check - neurosurgery}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_neurosurgery))
left_model <- with_dir(here(), tar_read(m_prior_left_neurosurgery))
right_model <- with_dir(here(), tar_read(m_prior_right_neurosurgery))
flat_model <- with_dir(here(), tar_read(m_prior_flat_neurosurgery))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

#### Ventriculostomy

```{r prior predictive check - evd}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_evd))
left_model <- with_dir(here(), tar_read(m_prior_left_evd))
right_model <- with_dir(here(), tar_read(m_prior_right_evd))
flat_model <- with_dir(here(), tar_read(m_prior_flat_evd))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

#### Tracheostomy

```{r prior predictive check - tracheostomy}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_tracheostomy))
left_model <- with_dir(here(), tar_read(m_prior_left_tracheostomy))
right_model <- with_dir(here(), tar_read(m_prior_right_tracheostomy))
flat_model <- with_dir(here(), tar_read(m_prior_flat_tracheostomy))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

#### Days of Mechanical Ventilation

```{r prior predictive check - vent}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_days_mechanical_ventilation))
left_model <- with_dir(here(), tar_read(m_prior_left_days_mechanical_ventilation))
right_model <- with_dir(here(), tar_read(m_prior_right_days_mechanical_ventilation))
flat_model <- with_dir(here(), tar_read(m_prior_flat_days_mechanical_ventilation))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

#### Comfort Care

```{r prior predictive check - comfort}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_comfort))
left_model <- with_dir(here(), tar_read(m_prior_left_comfort))
right_model <- with_dir(here(), tar_read(m_prior_right_comfort))
flat_model <- with_dir(here(), tar_read(m_prior_flat_comfort))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

#### Early WLST

```{r prior predictive check - early wlst}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_early_wlst))
left_model <- with_dir(here(), tar_read(m_prior_left_early_wlst))
right_model <- with_dir(here(), tar_read(m_prior_right_early_wlst))
flat_model <- with_dir(here(), tar_read(m_prior_flat_early_wlst))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

#### DNR

```{r prior predictive check - dnr}
#| echo: false
#| warning: false

neutral_model <- with_dir(here(), tar_read(m_prior_neutral_dnr_binary))
left_model <- with_dir(here(), tar_read(m_prior_left_dnr_binary))
right_model <- with_dir(here(), tar_read(m_prior_right_dnr_binary))
flat_model <- with_dir(here(), tar_read(m_prior_flat_dnr_binary))

results_visual(neutral_model)

f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)

```

## Posterior Simulation Diagnostics

```{r posterior diagnostics setup}
#| echo: false

posterior_diagnostics <- function(mod) {
  # plot code modified from Andrew Heiss (https://bayesf22-notebook.classes.andrewheiss.com/bayes-rules/13-chapter.html#simulating-the-posterior)

  trace_plot <- mod |>
    gather_draws(`^b_.*`, regex = TRUE) |>
    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
    geom_line(linewidth = 0.1) +
    scale_color_viridis_d(option = "rocket", end = 0.85) +
    facet_wrap(vars(.variable), scales = "free_y")

  trank_plot <- mod |>
    gather_draws(`^b_.*`, regex = TRUE) |>
    group_by(.variable) |>
    mutate(draw_rank = rank(.value)) |>
    ggplot(aes(x = draw_rank, color = factor(.chain))) +
    stat_bin(geom = "step", binwidth = 250, position = position_identity(), boundary = 0) +
    scale_color_viridis_d(option = "rocket", end = 0.85) +
    facet_wrap(vars(.variable), scales = "free_y") +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank())

  # pp_check <- pp_check(mod, ndraws = 500)
  # pp_check_bars <- pp_check(mod, ndraws = 500, type = "bars_grouped", group = "ich_laterality")

  return((trank_plot / trace_plot))
}

```

### MCMC Chain Convergence and Resolution

#### Neurosurgical Intervention

```{r posterior diagnostics neurosurgery}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_neurosurgery)))

```

#### Ventriculostomy

```{r posterior diagnostics evd}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_evd)))

```

#### Tracheostomy

```{r posterior diagnostics trach}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_tracheostomy)))

```

#### Days of Mechanical Ventilation

```{r posterior diagnostics vent}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)))

```

#### Comfort Care

```{r posterior diagnostics comfort}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_comfort)))

```

#### Early WLST

```{r posterior diagnostics early wlst}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_early_wlst)))

```

#### DNR

```{r posterior diagnostics dnr}
#| echo: false
#| fig-height: 10

posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_dnr_binary)))

```

### Posterior predictive checks

```{r posterior prediction function}
#| echo: false

post_pred <- function(mod) {
  pp_check <- pp_check(mod, ndraws = 500)

  pp_check_bars <- pp_check(mod, ndraws = 500, type = "bars_grouped", group = "ich_laterality")

  return(pp_check / pp_check_bars)
}

```

### Neurosurgical Intervention

```{r posterior predictions neurosurgery}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_neurosurgery)))

```

### Ventriculostomy

```{r posterior predictions evd}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_evd)))

```

### Tracheostomy

```{r posterior predictions trach}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_tracheostomy)))

```

### Days of Mechanical Ventilation

```{r posterior predictions vent}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)))

```

### Comfort Care

```{r posterior predictions comfort}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_comfort)))

```

### Early WLST

```{r posterior predictions early wlst}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_early_wlst)))

```

### DNR

```{r posterior predictions dnr}
#| echo: false

post_pred(with_dir(here(), tar_read(m_posterior_neutral_dnr_binary)))

```

## Sensitivity Analyses